
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part A
### Problem 1

The dataset consists of more than 2k+ job listing for data analyst`
positions. It has been scrapped from Glassdoor and gives insights into
the various job details like salary range, top companies, sectors, etc.

```{r}
library(readr)
library(tidyr)
```

```{r}
# importing data and replacing missing values with NA
df <- read_csv('G:/NEU_DS/IDMP/HW/2/2a/DataAnalyst.csv')
```

**Observations :** Missing values are represented by '-1' & 'Unknown'.

```{r}
# replacing -1 & Unknown by null value
df[df == "-1"] <- NA
df[df == "Unknown"] <- NA
```

```{r}
# checking dimension 
dim(df)

# checking total NA values in all columns
colSums(is.na(df))
```

**Observations :**

-   Column 1 has index no. and can be dropped.
-   Easy apply and Competitors have many missing values and can be dropped.
-   Job description, Headquarters, Founded can also be dropped as they
    are not required for our analysis.

```{r}
# printing column names to be dropped
col <- names(df[ ,c(1,4,8,10,15,16)])
col
```

```{r}
# dropping the above columns
df[,c(col)] <- list(NULL)
```

**Cleaning :**

-   Some job titles contain department name/ company name,etc. Most are
    separated from the title by a comma(,).
-   All company names have a integer value which need to be separated.
-   Salary Estimate column can be transformed to separate Min Salary &
    Max Salary columns and need to be converted to integer.

```{r}
library(stringr)
# splitting strings in columns 
df$`Job Title` <- str_split(df$`Job Title`, ",", simplify = TRUE)[,1]

df$`Company Name` <- str_split(df$`Company Name`, "\n", simplify = TRUE)[,1]

df$`Salary Estimate` <- str_split(df$`Salary Estimate`, " ", simplify = TRUE)[,1]
```

```{r}
# splitting Salary Estimate and dropping it
df$`Min Salary` <- str_split(df$`Salary Estimate`, "-", simplify = TRUE)[,1]
df$`Min Salary` <- as.integer(str_extract(df$`Min Salary`, "[[:digit:]]+"))

df$`Max Salary` <- str_split(df$`Salary Estimate`, "-", simplify = TRUE)[,2]
df$`Max Salary` <- as.integer(str_extract(df$`Max Salary`, "[[:digit:]]+"))

df$`Salary Estimate` <- NULL
```

All the data has been tidied and transformed.

### Problem 2

```{r}
library(ggplot2)

# Bar plot 
# Job distribution Ownership-wise
ggplot(df, aes(x=reorder(df$`Type of ownership`, df$`Type of ownership`, function(x)-length(x)))) +
  geom_bar(fill='blue') +  scale_y_log10() +
  labs(title="Job distribution Ownership-wise", x='Ownership Type') + coord_flip() +
  theme_classic()

# Job distribution Sector-wise
ggplot(df, aes(x=reorder(df$Sector, df$Sector, function(x)-length(x)))) +
  geom_bar(fill='green') + scale_y_log10() + 
  labs(title="Job distribution Sector-wise", x='Sectors') + coord_flip() +
  theme_classic()
```

**Observations :**

1)  The first plot shows the relationship between of jobs listings and
    the companies ownership type. 1k+ companies are privately backed
    followed by public companies. Rest are less than 100 in count.
2)  Jobs distribution according to different sectors are showcased
    above. Mining & Tourism sector have negligible listings while the
    healthcare, finance, business & IT sector have the majority
    openings.

```{r}
# Pie Chart
ggplot(data = df, aes(x = 1, fill=Size)) + 
  geom_bar() +
  coord_polar(theta="y") +
  labs(title = "Distribution of Employees Size", x= "", y="") +
  theme_classic()
```

**Observations :** The pie chart provides insights regarding the total number of employees
at companies. Most seem to have 200 employees followed by more than 1k.

```{r pressure}
# Box & Whisker plot
ggplot(df, aes(x=df$Sector, y=df$Rating)) + coord_flip() +
  geom_boxplot(fill='yellow') +
  labs(x='Sectors', y='Rating') +
  theme_classic()

ggplot(df, aes(x=((df$`Min Salary`+df$`Max Salary`)/2), y=df$`Type of ownership`)) +
  geom_boxplot(fill='orange') +
  labs(x='Average Salary', y='Ownership Type') +
  theme_classic()
```

**Observations :**

1)  The employee ratings w.r.t various sectors are plotted.
    Construction, Maintenance & Repair have low ratings while others are
    in the range 3-4.
2)  Average salary w.r.t Ownership type is plotted. Healthcare owned
    companies have the highest avg salary. The salary range is between
    60k-90k salary.

```{r}
# Histogram
ggplot(data=df, aes(x=df$`Min Salary`)) +
  geom_histogram(binwidth = 5) +  
  labs(title="Distribution of Minimum Salary", x="Minimum Salary (in 1000s)") +
  theme_classic()

ggplot(data=df, aes(x=df$`Max Salary`)) +
  geom_histogram(binwidth = 5) +  
  labs(title="Distribution of Maximum Salary", x="Maximum Salary (in 1000s)") +
  theme_classic()
```

**Observations:**

1)  The histogram shows the minimum/maximum salary for the Data Analyst
    positions of various companies. Both the distributions are right
    skewed.
2)  Most people have a minimum salary around 50k+. There are some
    outliers present where maximum salary is more than 150k+.

------------------------------------------------------------------------